---
title: "US Foreign Policy"
description: "Course syllabus"
date: "`r Sys.Date()`"
bibliography: ../../classrefs.bib
csl: bib/apsr-bibliography.csl
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

mon <- as.Date("2024-08-19")

advdate <- function(obj, adv){
  tmon <- obj + 7*(adv - 1)
  tfri <- obj + 4 + 7*(adv - 1)
  tmon <- format(tmon, format = "%m/%d")
  tfri <- format(tfri, format = "%m/%d")
  zadv <- sprintf("%02d", adv)
  tmp <- paste("Week ", zadv, sep = '', ", ", tmon, " - ", tfri)
  return(tmp)
}

```
 

## Course Summary
 
This course focuses on the development of US foreign policy over the course of the twentieth and twenty-first centuries. It will provide a brief background on the history of US foreign policy before 1900, but will focus primarily on the developments and issues occurring from 1900 through the present. The historical content of the course will be supplemented by introducing students to political science literature that can help to identify and explain key events and trends in US foreign policy during this time period. 

## Course Objectives

1. The overarching goal is to develop an understanding
of the systematic forces that shape US foreign policymaking. Please note that this is different than
prescribing what policy *should* be (though this is also an important consideration when considering our motivations for studying the topic and reading course materials.
2. Identify the key actors that influence foreign policy.
3. Understand how and when the influence of actors varies (e.g. across time, issue area, etc.).
4. Identify the key issues confronting policymakers and society. 
5. Identify general trends in US foreign policy over time. 
    - What's changed? 
    - What stays the same?
    - When and why do we see changes in behavior or policies?
6. Students should develop a familiarity with social science research processes, practices. Specifically, students should have:
    - A basic understanding of how to read social scientific research articles. 
    - A basic understanding of research design considerations
    - A basic understanding of what the scope and limitations of research articles


## Course Format 

The course will be part lecture and part student participation. I encourage students to ask questions if the readings of lecture materials are unclear. In addition to class lectures students will be responsible for keeping up with readings and other class-related material as the semester progresses.


## Required Books 

There are three required books and several required articles (listed below in the course schedule).  You can view the full citation in the reference list at the end of hover your cursor over the citation below for a preview of the full title.

- @Boczar2022
- @Busby2022
- @Doshi2023
- @Helberg2022

## Additional Readings

Students should also make an effort to stay informed on current events. Below are a some examples of good sources for keeping up with global events. Please note that some of these publications may be pay-walled, but you should have access through the university library.

  - [BBC News](www.bbc.com/news)
  - [Good Authority](https://goodauthority.org/)
  - [The Monkey Cage (No longer active)](www.washingtonpost.com/news/monkeycage)
  - [Political Violence @ a Glance](www.politicalviolenceataglance.org)
  - [Foreign Affairs](www.foreignaffairs.com)
  - [Foreign Policy Magazine](www.foreignpolicy.com)
  - [The Conversation](https://theconversation.com/us)
  - [The Washington Post](www.washingtonpost.com)
  - [The New York Times](www.nytimes.com)
  - [The Economist](www.economist.com)
  - [The Financial Times](www.ft.com)
  

## Course Requirements

1. _Attendance: 20%_

A major component to being successful in this class is just showing up. Every topic we cover will build upon ideas, lessons, themes, etc., that we covered in previous weeks. To do well you must be present.

2. _Collaborative Discussion: 20%_

While the course has a heavy lecture component students will also spend time reflecting on class materials and discussing that material with classmates. We will use a combination of small group discussions and a Think-Pair-Share format---Students will have an opportunity to reflect on the course material independently, discuss the material with a classmate, and then communicate a summation of that discussion with the rest of the class. Students will then submit the notes from their discussion for review and grading. Accordingly, use the notes to keep track of the content of the discussion, ideas that emerge, and any lingering questions that might arise.


3. _Content Quizzes: 20%_

There will be 6-8 content quizzes throughout the semester and they will cover material we have already reviewed through readings, lectures, and discussion. I will announce these ahead of time. All quizzes are open book and open note assignments and will be completed through Canvas. I will also identify the appropriate readings corresponding to each quiz so you do not have to worry about guessing. 


4. _DAG Exercise: 20%_

One of the things that separates the social sciences from punditry is the effort to be more precise in our understand of social phenomena. One tool that helps us to both crystallize our own thinking, and to communicate our thinking to others, is the directed acyclic graph (i.e. DAG). More casually, this is known as a causal graph. A DAG is simply a collection of nodes, which represent individual variables, and arrows, which represent causal ties. DAGs can range from relatively simple, to complex in structure, but they can all help us to identify causal relationships and develop model specifications to test those relationships.

Students should base their DAG exercise on a reading that we have covered *prior to the assignment*. Do not base it on future readings that we have not yet covered.

All DAG assignments will be completed in class. Students must complete the DAG exercise using pen/pencil and paper. Digital copies will not be accepted. 

The DAG assignments will become increasingly complex as we move through the course. We will spend the class period before each DAG due date reviewing some of the basic concepts associated with causality and drawing a DAG. We will begin with simplistic models and moving to more complex models. The goal is to help students to develop a better understanding of how to identify causal relationships and how to communicate those relationships to others.

The general pattern for the three assignments will be as follows:

1. DAG Exercise 1
  - Identify the key variables in the reading
  - Identify the causal relationships between the variables
  
2. DAG Exercise 2
  - Identify the key variables in the reading
  - Identify the causal relationships between the variables
  - Identify potential confounding variables

3. DAG Exercise 3
  - Identify the key variables in the reading
  - Identify the causal relationships between the variables
  - Identify potential confounding variables
  - Identify potential mediating variables
  - Identify potential moderating variables
  

5. _Photo Journal: 20%_

Students will be required to keep a photo journal You will be required to complete **3 entries** over the course of the semester. The goal of this assignment is to help students to develop a fuller appreciation for the ways in which their local community connects to the outside world. Here's the gist of the assignment:

  - Identify something around campus, in town, or in the nearby region that connects to US foreign policy and world politics. 
  - Take a picture of it!
  - Write journal entry in which you:
    - Share the picture
    - Tell us *what* it is
    - Tell us *where* it is
    - Explain *how* it ties into US foreign policy and world politics
    - Reference class materials to help support the connections and explain why it's important.

Be creative! There's really no right or wrong here, but it's on you to clearly explain how the subject of your photo ties into the course's subject matter. 

We will allocate one class period per month to completing the photo journal assignments. Students should submit a picture of their photo journal entry and a brief write-up of the entry. The write-up should be no more than 500 words. All write-ups will be completed in class using pen/pencil and paper. Digital copies will not be accepted.




  
## Classroom Policies

1. Be kind. This might seem like an odd thing to list first, but I want to emphasize that this is not a normal class environment---this is still a time of unparalleled stress for many of us. We're still in the middle of a global pandemic, many of us are still very much worried about our own health and safety, and many of us are very much worried about the health and safety of our coworkers, friends, and families. Let's try to be cognizant of the fact that many of us (if not all) are going to be struggling at various points, and let's work together to make this semester as educational and enjoyable as possible. 

2. Be professional. We're all here to learn, and many of the issues that we discuss have real-world impacts that affect people we know, or in our broader community. We're here to work together to develop a better understanding of these issues. 

3. Study guides and extra credit. The class is the study guide. I do not offer extra credit.

4. Grade Appeals. If you believe that I have given you an incorrect grade on an exam, you may submit
a written appeal. All appeals must be in writing (they may not be made in person) and must contain
an explanation for why the grade is incorrect. I will also require you to wait 24 hours after receiving a
grade to submit a written appeal. Please note that even if your appeal is granted, I reserve the right
to regrade the assignment in its entirety, meaning that your grade may go up or down depending on
my reevaluation of the assignment.

5. The Syllabus. The syllabus is a living document that can and will be altered throughout the duration
of the course based both on need and design. Generally, this may mean readings will be removed or
added as needed. All changes will be listed on K-State Online, so make sure to check announcements
for any such change. Additionally, I maintain the right to all of my own intellectual property presented
in this course, whether it is the course lectures or this syllabus. Materials from this course ought not
to be reproduced without my expressed permission.

6. Classroom conduct. All student activities in the University, including this course, are governed by
the Student Judicial Conduct Code as outlined in the Student Governing Association By Laws, Article
V, Section 3, number 2. Students who engage in behavior that disrupts the learning environment may
be asked to leave the class.
  
  


## Contacting the Instructor

The best way to get in touch with me is through email. Please do not hesitate to contact me if you have any questions about course content or your ability to complete the assigned material.



## Final Grades

A: 90+

B: 80–89.99

C: 70–79.99

D: 60–69.99

F: 0–59.99


*Tentative course schedule begins on next page*

\clearpage


## Tentative Course Schedule

```{r course calendar, dpi = 300, warning = FALSE, echo = FALSE, message = FALSE, fig.height=9, fig.width=7, fig.pos="center"}

pacman::p_load(
  "ggplot2",
  "showtext"
)

# Add custom fonts
#sysfonts::font_add_google("Roboto", family = "oswald")
#showtext::showtext_auto()

# Create base data frame
dfr <- data.frame(date=seq(as.Date('2024-08-01'), as.Date('2024-12-31'), by = 1))
dfr$day <- factor(strftime(dfr$date,format="%a"), levels=c("Sun","Mon","Tue","Wed","Thu","Fri","Sat"))
dfr$week <- factor(strftime(dfr$date,format="%U"))
dfr$month <- factor(strftime(dfr$date,format="%B"),levels=c("August","September","October", "November", "December"))
dfr$ddate <- factor(strftime(dfr$date,format="%d"))



# Photo blogs
dfr$comment[dfr$date==as.Date("2024-09-27")] <- "Photo Journal Day"
dfr$comment[dfr$date==as.Date("2024-10-25")] <- "Photo Journal Day"
dfr$comment[dfr$date==as.Date("2024-11-22")] <- "Photo Journal Day"

# DAG Exercises
dfr$comment[dfr$date==as.Date("2024-09-13")] <- "DAG Exercise"
dfr$comment[dfr$date==as.Date("2024-10-11")] <- "DAG Exercise"
dfr$comment[dfr$date==as.Date("2024-11-08")] <- "DAG Exercise"

# Book Discussion
dfr$comment[dfr$date==as.Date("2024-10-18")] <- "Boczar Discussion"
dfr$comment[dfr$date==as.Date("2024-11-15")] <- "Busby Discussion"
dfr$comment[dfr$date==as.Date("2024-11-01")] <- "Helberg Discussion"
dfr$comment[dfr$date==as.Date("2024-12-06")] <- "Doshi Discussion"


# Thanksgiving
dfr$comment[dfr$date >= as.Date("2024-11-25") & dfr$date <= as.Date("2024-11-29")] <- "Thanksgiving Break"

dfr$week <- as.numeric(dfr$week) - min(as.numeric(dfr$week) - 1)


ggplot(dfr, aes(x = day, y = reorder(week, -week), fill = comment)) +
  geom_tile(color = "black", size = 0.1) +
  geom_text(aes(label = ddate), size = 3) +
  facet_wrap(. ~ month, scales = "free", ncol = 2) +
  theme_void() +
  theme(legend.position = "bottom",
        legend.direction = "vertical",
        legend.justification = "right",
        axis.text.x = element_text(size = 10, color = "black"),
        axis.text.y = element_text(size = 10, color = "white"),
        strip.text.x = element_text(size = 13, face = "bold", hjust = 0),
        strip.text = element_text(size = 16, face = "bold", margin = margin(b = 0.2, unit = "cm")),
        title = element_text(face = "bold"),
        panel.spacing = unit(1, "cm")) +
  guides(fill=guide_legend(ncol=2)) +
  scale_fill_brewer(palette = "Set2", na.translate = FALSE) +
  labs(x = "Weekday",
       y = "Week #",
       fill = "Assignment")
```
  
  
  
\clearpage


## `r advdate(mon, 1)`: Course Introduction

### Readings and Assignments

1. Watch video on [how to read a social science research paper](https://www.youtube.com/watch?v=pZq6c_shco0)


## `r advdate(mon, 2)`: Historical background

### Readings and Assignments: 

- Read @Fordham2019



## `r advdate(mon, 3)`: The Cold War and US Hegemony

### Readings and Assignments:

- Read @NSC68



## `r advdate(mon, 4)`: The Cold War and Domestic Politics

### Readings and Assignments:

- Read @Dudziak2004
- Read @Fordham2007


## `r advdate(mon, 5)`: The Presidency and Congress

### Readings and Assignments:

- Read @Lupton2017


## `r advdate(mon, 6)`: The Policymaking Process

### Readings and Assignments:

- Read @Miller2016



## `r advdate(mon, 7)`: The Bureaucracy / The State Department

### Readings and Assignments:

- Wednesday September 25: Read @KeaneDiesen2015



## `r advdate(mon, 8)`: The Defense Department and the Military

### Readings and Assignments:




## `r advdate(mon, 9)`: The Use of Military Force

### Readings and Assignments:


- Read @Boczar2022

Suggested:

- Watch [Restrepo](https://watchdocumentaries.com/restrepo/)
- Read @sigar2019



## `r advdate(mon, 10)`: Security Cooperation

### Readings and Assignments:


Wednesday, October 23 Read:
- [Flynn, Allen, and Martinez Machain 2019](https://theconversation.com/why-does-the-us-pay-so-much-for-the-defense-of-its-allies-5-questions-answered-127683)
 
- [Allen, Martinez Machain, and Flynn 2022](https://theconversation.com/the-us-military-presence-in-europe-has-been-declining-for-30-years-the-current-crisis-in-ukraine-may-reverse-that-trend-175595)


## `r advdate(mon, 11)`: Trade, Immigration, Finance, and Monetary Policy

### Readings and Assignments:

- Read @Helberg2022
- Read @Oneil2022


## `r advdate(mon, 12)`: Trade, Immigration, Finance, and Monetary Policy (cont.)

### Readings and Assignments:





## `r advdate(mon, 13)`: Environmental Politics and US Foreign Policy

- Read @Busby2022




## `r advdate(mon, 14)`: Human Rights Issues

### Readings and Assignments:

- Read @Ballve2009
- Read @Diamond2023



## `r advdate(mon, 15)`: Thanksgiving Break (No Class)

### Readings and Assignments:

1. Get some rest
2. Enjoy yourselves
3. Eat good food
4. Take a good nap


## `r advdate(mon, 16)`: US and Chinese Competition

### Readings and Assignments:

- Read @Doshi2023





 

