---
title: "US Foreign Policy"
description: "Course syllabus"
date: "`r Sys.Date()`"
bibliography: ../../classrefs.bib
csl: bib/apsr-bibliography.csl
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

mon <- as.Date("2023-08-21")

advdate <- function(obj, adv){
  tmon <- obj + 7*(adv - 1)
  tfri <- obj + 4 + 7*(adv - 1)
  tmon <- format(tmon, format = "%m/%d")
  tfri <- format(tfri, format = "%m/%d")
  zadv <- sprintf("%02d", adv)
  tmp <- paste("Week ", zadv, sep = '', ", ", tmon, " - ", tfri)
  return(tmp)
}

```
 

## Course Summary
 
This course focuses on the development of US foreign policy over the course of the twentieth and twenty-first centuries. It will provide a brief background on the history of US foreign policy before 1900, but will focus primarily on the developments and issues occurring from 1900 through the present. The historical content of the course will be supplemented by introducing students to political science literature that can help to identify and explain key events and trends in US foreign policy during this time period. 

## Course Objectives

1. The overarching goal is to develop an understanding
of the systematic forces that shape US foreign policymaking. Please note that this is different than
prescribing what policy *should* be (though this is also an important consideration when considering our motivations for studying the topic and reading course materials.
2. Identify the key actors that influence foreign policy.
3. Understand how and when the influence of actors varies (e.g. across time, issue area, etc.).
4. Identify the key issues confronting policymakers and society. 
5. Identify general trends in US foreign policy over time. 
    - What's changed? 
    - What stays the same?
    - When and why do we see changes in behavior or policies?
6. Students should develop a familiarity with social science research processes, practices. Specifically, students should have:
    - A basic understanding of how to read social scientific research articles. 
    - A basic understanding of research design considerations
    - A basic understanding of what the scope and limitations of research articles


## Course Format 

The course will be part lecture and part student participation. I encourage students to ask questions if the readings of lecture materials are unclear. In addition to class lectures students will be responsible for keeping up with readings and other class-related material as the semester progresses.


## Required Books 

There are three required books and several required articles (listed below in the course schedule).  You can view the full citation in the reference list at the end of hover your cursor over the citation below for a preview of the full title.

- @Allenetal2022
- @GeorgeRishikof2017
- @Karlin2022
- @McDowell2023

## Additional Readings

Students should also make an effort to stay informed on current events. Below are a some examples of good sources for keeping up with global events. Please note that some of these publications may be pay-walled, but you should have access through the university library.

  - [BBC News](www.bbc.com/news)
  - [The Monkey Cage](www.washingtonpost.com/news/monkeycage)
  - [Political Violence @ a Glance](www.politicalviolenceataglance.org)
  - [Foreign Affairs](www.foreignaffairs.com)
  - [Foreign Policy Magazine](www.foreignpolicy.com)
  - [The Conversation](https://theconversation.com/us)
  - [The Washington Post](www.washingtonpost.com)
  - [The New York Times](www.nytimes.com)
  - [The Economist](www.economist.com)
  - [The Financial Times](www.ft.com)
  

## Course Requirements

1. _Collaborative Discussion: 30%_

While the course has a heavy lecture component students will also spend time reflecting on class materials and discussing that material with classmates. We will use a combination of small group discussions and a Think-Pair-Share format---Students will have an opportunity to reflect on the course material independently, discuss the material with a classmate, and then communicate a summation of that discussion with the rest of the class. Students will then submit the notes from their discussion for review and grading. Accordingly, use the notes to keep track of the content of the discussion, ideas that emerge, and any lingering questions that might arise.


2. _Content Quizzes: 20%_

There will be 6-8 content quizzes throughout the semester and they will cover material we have already reviewed through readings, lectures, and discussion. I will announce these ahead of time. All quizzes are open book and open note assignments and will be completed through Canvas. I will also identify the appropriate readings corresponding to each quiz so you do not have to worry about guessing. 


3. _DAG Exercise: 20%_

One of the things that separates the social sciences from punditry is the effort to be more precise in our understand of social phenomena. One tool that helps us to both crystalize our own thinking, and to communicate our thinking to others, is the directed acyclic graph (i.e. DAG). More casually, this is known as a causal graph. A DAG is simply a collection of nodes, which represent individual variables, and arrows, which represent causal ties. DAGs can range from relatively simple, to complex in structure, but they can all help us to identify causal relationships and develop model specifications to test those relationships.

Students should choose three readings to serve as a basis for their DAG assignment. These readings should also be from separate weeks and NOT from the same week. Rather, don't wait to the last week of class to cram in three DAG assignments.

Each DAG assignment will consist of the following parts:

1. **Question:** After reading one of our in-class reading assignments (article, book, etc.) students will identify a new research question.
2. **Objective:** What is the objective of the research question? Students should clearly describe the goal and the scope of the proposed question/analysis.
3. **Estimand:** Borrowing from @Lundbergetal2021, the estimand consists of 1) a unit-specific quantity, and 2) a description of the target population. Importantly, you must explain how this estimand relates to your research question. Is it an appropriate measure given your question?
4. **DAG:** Using the [Dagitty web platform](https://www.dagitty.net/dags.html)


  
4. _The Music of Foreign Policy Wiki: 30%_

The arts are a powerful tool for communicating ideas across space and time, and are often deeply rooted in the politics of the moment. Over the last century music has increasingly come to serve as a vehicle for political speech, both reflecting and driving politics in the modern era. To that end, students will work together in small groups to build a Wiki-style page presenting a song or album that relates to US foreign policy in some way. The goal of this project is for students and their groups to build a wiki page on Canvas is dedicated to providing readers with a deeper understanding of the historical, political, and social underpinnings of songs. But there's a catch—since this is a class on US foreign policy, these songs have to be related to US foreign policy in some way. 

In general, each group's wiki page should focus on the following points:

  - *Song/Album*: Identify a song (or if you're feeling ambitious, an entire album) that speaks to some aspect of US foreign policy.
  - *Issue*: Identify the central issue, actors, policy, action, idea, behavior, etc., that is at the core of the song. 
  - *Context*: Once you have identified the central issue, provide some historical background information. This should help readers to better understand the social, political, and economic issues of the time period in question. Who are the key actors? What are the issues or policies of the time that serve as the focus for the song/album? 
  - *Reception*: How was the song/album received at the time of its release? Both critically and with respect to the popular impact that it had at the time?
  - *Legacy*: What is the legacy of the song/album? How does its legacy compare with its impact at the time of its release?


This will serve as your final project. Accordingly, I expect the final content to be detailed, polished, and thoroughly proofed. Your grade will be broken up into components and there will be multiple checkpoints along the way. The schedule is detailed below.

Deadlines: 

- September 9: Song/album choices due (2.5%)
- September 23: Final song/album selection due (2.5%)
- October 28: Annotated bibliography due (10%)
- November 18: Wiki draft due  (5%)
- December 9: Final Wiki page due (10%)
  
Each iteration of this assignment should be drafted on Microsoft Word or Apple Pages, and submitted to Canvas.
  
## Classroom Policies

1. Be kind. This might seem like an odd thing to list first, but I want to emphasize that this is not a normal class environment---this is still a time of unparalleled stress for many of us. We're still in the middle of a global pandemic, many of us are still very much worried about our own health and safety, and many of us are very much worried about the health and safety of our coworkers, friends, and families. Let's try to be cognizant of the fact that many of us (if not all) are going to be struggling at various points, and let's work together to make this semester as educational and enjoyable as possible. 

2. Be professional. We're all here to learn, and many of the issues that we discuss have real-world impacts that affect people we know, or in our broader community. We're here to work together to develop a better understanding of these issues. 

3. Study guides and extra credit. The class is the study guide. I do not offer extra credit.

4. Grade Appeals. If you believe that I have given you an incorrect grade on an exam, you may submit
a written appeal. All appeals must be in writing (they may not be made in person) and must contain
an explanation for why the grade is incorrect. I will also require you to wait 24 hours after receiving a
grade to submit a written appeal. Please note that even if your appeal is granted, I reserve the right
to regrade the assignment in its entirety, meaning that your grade may go up or down depending on
my reevaluation of the assignment.

5. The Syllabus. The syllabus is a living document that can and will be altered throughout the duration
of the course based both on need and design. Generally, this may mean readings will be removed or
added as needed. All changes will be listed on K-State Online, so make sure to check announcements
for any such change. Additionally, I maintain the right to all of my own intellectual property presented
in this course, whether it is the course lectures or this syllabus. Materials from this course ought not
to be reproduced without my expressed permission.

6. Classroom conduct. All student activities in the University, including this course, are governed by
the Student Judicial Conduct Code as outlined in the Student Governing Association By Laws, Article
V, Section 3, number 2. Students who engage in behavior that disrupts the learning environment may
be asked to leave the class.
  
  


## Contacting the Instructor

The best way to get in touch with me is through email. Please do not hesitate to contact me if you have any questions about course content or your ability to complete the assigned material.



## Final Grades

A: 90+

B: 80–89.99

C: 70–79.99

D: 60–69.99

F: 0–59.99


*Tentative course schedule begins on next page*

\clearpage


## Tentative Course Schedule

```{r course calendar, dpi = 300, warning = FALSE, echo = FALSE, message = FALSE, fig.height=9, fig.width=7, fig.pos="center"}

pacman::p_load(
  "ggplot2",
  "showtext"
)

# Add custom fonts
#sysfonts::font_add_google("Roboto", family = "oswald")
#showtext::showtext_auto()

# Create base data frame
dfr <- data.frame(date=seq(as.Date('2023-08-01'), as.Date('2023-12-31'), by = 1))
dfr$day <- factor(strftime(dfr$date,format="%a"), levels=c("Sun","Mon","Tue","Wed","Thu","Fri","Sat"))
dfr$week <- factor(strftime(dfr$date,format="%U"))
dfr$month <- factor(strftime(dfr$date,format="%B"),levels=c("August","September","October", "November", "December"))
dfr$ddate <- factor(strftime(dfr$date,format="%d"))

# Label dates
dfr$comment <- "Available"
dfr$comment[dfr$date < as.Date("2023-08-21") | dfr$date>as.Date("2021-12-09")] <- NA
dfr$comment[dfr$date==as.Date("2023-09-08")] <- "Song/album choice due"
dfr$comment[dfr$date==as.Date("2023-09-22")] <- "Final song/album choice due"
dfr$comment[dfr$date==as.Date("2023-10-27")] <- "Annotated bibliography due"
dfr$comment[dfr$date==as.Date("2023-11-17")] <- "Wiki draft due"
dfr$comment[dfr$date==as.Date("2023-12-08")] <- "Final Wiki due"

# Photo blogs
dfr$comment[dfr$date==as.Date("2023-09-30")] <- "Blog post due"
dfr$comment[dfr$date==as.Date("2023-10-31")] <- "Blog post due"
dfr$comment[dfr$date==as.Date("2023-11-30")] <- "Blog post due"


# Thanksgiving
dfr$comment[dfr$date >= as.Date("2023-11-20") & dfr$date <= as.Date("2023-11-24")] <- "Thanksgiving Break"

dfr$week <- as.numeric(dfr$week) - min(as.numeric(dfr$week) - 1)
dfr$comment <- factor(dfr$comment, levels=c("Blog post due", 
                                            "Song/album choice due", 
                                            "Final song/album choice due", 
                                            "Annotated bibliography due", 
                                            "Wiki draft due", 
                                            "Thanksgiving Break", 
                                            "Final Wiki due"))

ggplot(dfr, aes(x = day, y = reorder(week, -week), fill = comment)) +
  geom_tile(color = "black", size = 0.1) +
  geom_text(aes(label = ddate), size = 3) +
  facet_wrap(. ~ month, scales = "free", ncol = 2) +
  theme_void() +
  theme(legend.position = "bottom",
        legend.direction = "vertical",
        legend.justification = "right",
        axis.text.x = element_text(size = 10, color = "black"),
        axis.text.y = element_text(size = 10, color = "white"),
        strip.text.x = element_text(size = 13, face = "bold", hjust = 0),
        strip.text = element_text(size = 16, face = "bold", margin = margin(b = 0.2, unit = "cm")),
        title = element_text(face = "bold")) +
  guides(fill=guide_legend(ncol=2)) +
  scale_fill_brewer(palette = "Set2", na.translate = FALSE) +
  labs(x = "Weekday",
       y = "Week #",
       fill = "Assignment")
```
  
  
  
\clearpage


## `r advdate(mon, 1)`: Course Introduction

### Readings and Assignments

1. Watch video on [how to read a social science research paper](https://www.youtube.com/watch?v=pZq6c_shco0)


## `r advdate(mon, 2)`: Historical background

### Readings and Assignments: 

- Read @Fordham2019



## `r advdate(mon, 3)`: The Cold War and US Hegemony

### Readings and Assignments:

- Read @NSC68



## `r advdate(mon, 4)`: The Cold War and Domestic Politics

### Readings and Assignments:

- Read @Dudziak2004
- Read @Fordham2007


## `r advdate(mon, 5)`: The Presidency and Congress

### Readings and Assignments:

- Read @GeorgeRishikof2017[Chapter 13]
- Read @Lupton2017


## `r advdate(mon, 6)`: The Policymaking Process

### Readings and Assignments:

- Read @GeorgeRishikof2017[Chapters 1 and 2]
- Read @Miller2016



## `r advdate(mon, 7)`: The Bureaucracy / The State Department

### Readings and Assignments:

- Read @KeaneDiesen2015
- Read @GeorgeRishikof2017[Chapter 4]



## `r advdate(mon, 8)`: The Defense Department and the Military

### Readings and Assignments:

- Read @Karlin2022



## `r advdate(mon, 9)`: The Use of Military Force

### Readings and Assignments:

- Watch [Restrepo](https://watchdocumentaries.com/restrepo/)
- Read @sigar2019



## `r advdate(mon, 10)`: Security Cooperation

### Readings and Assignments:

- Read @Allenetal2022


## `r advdate(mon, 11)`: Trade, Immigration, Finance, and Monetary Policy

### Readings and Assignments:

- Read @Oneil2022


## `r advdate(mon, 12)`: Trade, Immigration, Finance, and Monetary Policy (cont.)

### Readings and Assignments:

- Read @McDowell2023



## `r advdate(mon, 13)`: Foreign Aid

- Read @GeorgeRishikof2017[Chapter 5]
- Read @PEPFARat20
- Read @Fidler2023


## `r advdate(mon, 14)`: Thanksgiving Break (No Class)

### Readings and Assignments:

1. Get some rest
2. Enjoy yourselves
3. Eat good food
4. Take a good nap


## `r advdate(mon, 15)`: Human Rights Issues

### Readings and Assignments:

- Read @Fajardo-Heyward2015
- Read @Ballve2009
- Read @Diamond2023


## `r advdate(mon, 16)`: Multilateralism and global governance

### Readings and Assignments:

- Read @Schultz2017
- Read @Busby2020 





 

