---
title: "World Politics"
description: "Course syllabus"
date: "`r Sys.Date()`"
bibliography: ../../classrefs.bib
csl: bib/apsr-bibliography.csl
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

mon <- as.Date("2023-01-16")

advdate <- function(obj, adv){
  tmon <- obj + 7*(adv - 1)
  tfri <- obj + 4 + 7*(adv - 1)
  tmon <- format(tmon, format = "%m/%d")
  tfri <- format(tfri, format = "%m/%d")
  zadv <- sprintf("%02d", adv)
  tmp <- paste("Week ", zadv, sep = '', ", ", tmon, " -- ", tfri)
  return(tmp)
}

```
 

## Course Summary
 
This course is designed to give undergraduate students an introduction to the topics and methods associated with the study of world politics and international relations. Over the past several centuries the type and frequency of interactions between peoples and governments have increased exponentially. These interactions range from the peaceful---trade, investment, athletics, tourism, study abroad, etc.---to the conflictual and tragic---famine, pandemics, human rights violations, and more. Furthermore, the proliferation of communications technology has allowed us to watch events unfold. in real time from almost anywhere on the globe. 

Given that such events---whether social, political, or economic---are increasingly relevant for individuals around the world, it is imperative that we develop a solid foundation for understanding the actors, incentives, and processes that drive them. Accordingly, this course will focus on basic facts and key concepts related to the study of international relations. We will also focus on how to develop mental models that can help us to understand broader patterns of behavior in world affairs, and to analyze the casual processes underpinning events of interest.


## Course Objectives

1. Develop an understanding of some of the key issues in world politics.
2. Identify some of the most important actors as they relate to these key issues.
3. Develop an understanding of evidence-based research as it relates to these issues and actors:
   - Develop an understanding of the relationship between observation, theory, evidence, and analysis.
   - Identify questions, problems, or puzzles in world politics.
   - Think about how we can develop theoretical models to help us analyze these issues.
   - Think about how we can structure research to analyze these questions.
   - Think about how we can collect different types of data to help us answer these questions.
   - Think about how we can evaluate evidence to draw conclusions and generate insights.
4. Learn to think about issues in world politics in terms of correlation, causation, and evidence.
  
  

## Course Format 

The course will be primarily lecture, thought students are encouraged to participate and ask questions where appropriate. I encourage students to ask questions if the readings of lecture materials are unclear and to reach out if they need help. In addition to class lectures students will be responsible for keeping up with readings and other class-related material as the semester progresses.


## Required Readings 

@FLS2021




## Additional Readings

Students should also make an effort to stay informed on current events. Below are a some examples of good sources for keeping up with global events. Please note that some of these publications may be pay-walled, but you should have access through the university library.

  - [BBC News](www.bbc.com/news)
  - [The Monkey Cage](www.washingtonpost.com/news/monkeycage)
  - [Political Violence @ a Glance](www.politicalviolenceataglance.org)
  - [Foreign Affairs](www.foreignaffairs.com)
  - [Foreign Policy Magazine](www.foreignpolicy.com)
  - [The Washington Post](www.washingtonpost.com)
  - [The New York Times](www.nytimes.com)
  - [The Economist](www.economist.com)
  - [The Financial Times](www.ft.com)
  

## Course Requirements


1. **Exams: 70%**:
Your grade will largely be determined by a series of four tests. These are designed to assess your knowledge of basic ideas and concepts that we will cover throughout the course. All tests will be open book/note, and will be administered online. They will cover material from the text book, in-class
lectures, discussions, podcasts, and/or any additional assigned readings. For each test you will have 75 minutes to answer $\approx$ 35 questions. There are no separate mid-term or final exams.

2. **Homework: 30%**:
You will have several homework assignments that are designed to provide you with deeper exposure to the course concepts. There will be approximately 6 homework assignments throughout the semester. All require students to complete various exercises and submit their work to the instructor. Some of these exercises will be completed outside of class and will be turned in through Canvas. Other activities will be completed in class and will be submitted to the instructor at the end of the class session. 




## Course Policies

1. **Be kind**. The past three years have been a time of unparalleled stress for many of us. We're still in the middle of a global pandemic, many of us are still very much worried about our own health and safety, and many of us are very much worried about the health and safety of our coworkers, friends, and families. Zoom fatigue is also a thing, so please be patient with one another. Let's try to be cognizant of the fact that some of us (if not all) may still be struggling at various points, and let's work together to make this semester as educational and enjoyable as possible. 

2. **Be professional**. Please try to treat your interactions with everyone as professionally as you would in a workplace environment.   

3. **Study guides and extra credit**. The class *is* the study guide. I do not offer extra credit.

4. **Grade Appeals**. If you believe that I have given you an incorrect grade on an exam, you may submit a written appeal. All appeals must be in writing (they may not be made in person) and must contain an explanation for why the grade is incorrect. I will also require you to wait 24 hours after receiving a
grade to submit a written appeal. Please note that even if your appeal is granted, I reserve the right to regrade the assignment in its entirety, meaning that your grade may go up or down depending on my reevaluation of the assignment.

5. **Missing Assignments and Makeup Work**. I will allow students to make up work that they miss do to a University-excused absence or event. In such cases students should provide two weeks notice and documentation provided by The Office of Student Life.  Students who miss assignments due to illness may make up assignments provided they have documentation from Student Health Services.

6. **The Syllabus**. The syllabus is a living document that can and will be altered throughout the duration of the course based both on need and design. Generally, this may mean readings will be removed or added as needed. All changes will be listed on K-State Online, so make sure to check announcements for any such change. Additionally, I maintain the right to all of my own intellectual property presented in this course, whether it is the course lectures or this syllabus. Materials from this course ought not to be reproduced without my expressed permission.

7. **Contacting the Instructor**. The best way to contact me is via email. I will usually respond within two business days, but if it takes a little longer do not hesitate to send a reminder email. I do not respond to emails over the weekend. 

    If you wish to make an appointment you can use the Calendly link below to find an available time. Given how hectic everyone's schedules are I will assume that these meetings will be Zoom based unless otherwise requested. 

    Email: [meflynn@ksu.edu](meflynn@ksu.edu)
    
    Appointments: [Make an appointment here](https://calendly.com/meflynn)


## Statement for Copyright Notification

Copyright 2023 (Michael Flynn) as to this syllabus and all lectures. During this course students are prohibited from selling notes to or being paid for taking notes by any person or commercial firm without the express written permission of the professor teaching this course. In addition, students in this class are not authorized to provide class notes or other class-related materials to any other person or entity, other than sharing them directly with another student taking the class for purposes of studying, without prior written permission from the professor teaching this course.


## Citation and Style

All written work is expected to adhere to professional standards. As described above, citing previous work is key. If you have any questions regarding the appropriate citation style please consult the American Political
Science Association style guide.




## University Syllabi Statements
  
  
### Academic Honesty and Plagiarism
  
All work turned in as a part of this class must be original work by the student.

Kansas State University has an Honor and Integrity System based on personal integrity, which is presumed to be sufficient assurance that, in academic matters, one's work is performed honestly and without unauthorized assistance. Undergraduate and graduate students, by registration, acknowledge the jurisdiction of the Honor and Integrity System. The policies and procedures of the Honor and Integrity System apply to all full and part-time students enrolled in undergraduate and graduate courses on-campus, off-campus, and via distance learning. A component vital to the Honor and Integrity System is the inclusion of the Honor Pledge which applies to all assignments, examinations, or other course work undertaken by students. The Honor Pledge is implied, whether or not it is stated: "On my honor, as a student, I have neither given nor received unauthorized aid on this academic work." A grade of XF can result from a breach of academic honesty. The F indicates failure in the course; the X indicates the reason is an Honor Pledge violation.



For more information, visit the [Honor & Integrity System home web page](http://www.k-state.edu/honor/).
All work is expected to adhere to the standards outlined by the Kansas State University Handbook. All written work is subject to online plagiarism checks and will be required for all reaction papers as well as the final paper. Any cases of plagiarism will receive an automatic failure for the assignment, possible failure in the course, and will be pursued further through the university for additional punishment.

The Kansas State University Handbook (2008, Appendix F) defines plagiarism as follows:

> Plagiarism is taking credit for someone else’s ideas, work, or words. In a university setting, it
means submitting academic, scholarly, or literary work in which you either claim or imply the
material to be your own, when that is not the case.

The Handbook also provides examples of plagiarism, and steps students can take to avoid it:

[P]lagiarism could include:

  - Buying a term paper, or copying another person’s paper, even if she or he gives you permission.
  - Cutting and pasting information from the internet into your own paper without properly
identifying it as a quotation and properly attributing the source.
  - Using someone else’s concept and presenting it as if it is your own original creativity, without
acknowledging the source of the idea.

While some acts of plagiarism are obvious attempts to deceive, like buying a term paper, others result from sloppy scholarship or failure to follow proper format for crediting sources. For example:

  - If you copy directly from another source and acknowledge it in your bibliography or list of works cited, yet fail to put it in quotation marks, this is plagiarism. The reason is that, although you have indicated that the idea is someone else’s, the lack of quotation marks
implies that the words are your own.
  - Paraphrasing is summarizing a source so that you have rewritten the material in your own words but maintained the original author’s ideas. When you paraphrase, you do not need to put the words in quotation marks, but you do need to properly attribute the original source. Paraphrasing does not mean simply changing one or two words, or leaving out a sentence, while the rest remains the same. Even if you identify the original author, but have used faulty paraphrasing that retains too much of the original, the result is inadvertent plagiarism.
  - If you use another researcher’s specific methodology, you should acknowledge doing so. While some research protocols are widespread and general enough to need no citation, if the protocol, analysis, or technique can be attributed to a specific publication it should be cited.
  - Works of visual art and design, and images of the work, cannot be misrepresented as original work by another. Such images cannot be used in part or whole without permission of the artist.

Also note that using papers previously written for another assignment in this, or in any other class, without the express permission of the professor is grounds for failing an assignment. Using your own material in this way, and not citing it, is also technically plagiarism and can be treated as such. If you have concerns regarding any of these issues please do not hesitate to contact me.


### Students with Disabilities

At K-State it is important that every student has access to course content and the means to demonstrate course mastery. Students with disabilities may benefit from services including accommodations provided by the Student Access Center. Disabilities can include physical, learning, executive functions, and mental health. You may register at the Student Access Center or to learn more contact:  

Manhattan/Olathe/Global Campus – Student Access Center  
accesscenter@k-state.edu
785-532-6441    

K-State Salina Campus – Julie Rowe; Student Success Coordinator
jarowe@k-state.edu
785-820-7908    

Students already registered with the Student Access Center please request your Letters of Accommodation early in the semester to provide adequate time to arrange your approved academic accommodations. Once SAC approves your Letter of Accommodation it will be e-mailed to you, and your instructor(s) for this course.  Please follow up with your instructor to discuss how best to implement the approved accommodations. 


### Classroom conduct 

All student activities in the University, including this course, are governed by the Student Judicial Conduct Code as outlined in the Student Governing Association By Laws, Article V, Section 3, number 2. Students who engage in behavior that disrupts the learning environment may be asked to leave the class.


### Statement on Mutual Respect and Inclusion in K-State Teaching and Learning Spaces

At K-State, faculty and staff are committed to creating and maintaining an inclusive and supportive learning environment for students from diverse backgrounds and perspectives. K-State courses, labs, and other virtual and physical learning spaces promote equitable opportunity to learn, participate, contribute, and succeed, regardless of age, race, color, ethnicity, nationality, genetic information, ancestry, disability, socioeconomic status, military or veteran status, immigration status, Indigenous identity, gender identity, gender expression, sexuality, religion, culture, as well as other social identities.

Faculty and staff are committed to promoting equity and believe the success of an inclusive learning environment relies on the participation, support, and understanding of all students. Students are encouraged to share their views and lived experiences as they relate to the course or their course experience, while recognizing they are doing so in a learning environment in which all are expected to engage with respect to honor the rights, safety, and dignity of others in keeping with the K-State Principles of Community. 

If you feel uncomfortable because of comments or behavior encountered in this class, you may bring it to the attention of your instructor, advisors, and/or mentors. If you have questions about how to proceed with a confidential process to resolve concerns, please contact the Student Ombudsperson Office. Violations of the student code of conduct can be reported using the Code of Conduct Reporting Form. You can also report discrimination, harassment or sexual harassment, if needed.



### Statement Regarding Wearing of Face Coverings

Kansas State University strongly encourages, but does not require, that everyone wear masks while indoors on university property, including while attending in-person classes. For additional information and the latest updates, see K-State’s face covering policy.



### Statement Regarding Discrimination, Harassment, and Sexual Harassment

Kansas State University is committed to maintaining academic, housing, and work environments that are free of discrimination, harassment, and sexual harassment. Instructors support the University’s commitment by creating a safe learning environment during this course, free of conduct that would interfere with your academic opportunities. Instructors also have a duty to report any behavior they become aware of that potentially violates the University’s policy prohibiting discrimination, harassment, and sexual harassment, as outlined by PPM 3010.  

If a student is subjected to discrimination, harassment, or sexual harassment, they are encouraged to make a non-confidential report to the University’s Office for Institutional Equity (OIE) using the online reporting form. Incident disclosure is not required to receive resources at K-State. Reports that include domestic and dating violence, sexual assault, or stalking, should be considered for reporting by the complainant to the Kansas State University Police Department or the Riley County Police Department. Reports made to law enforcement are separate from reports made to OIE. A complainant can choose to report to one or both entities. Confidential support and advocacy can be found with the K-State Center for Advocacy, Response, and Education (CARE). Confidential mental health services can be found with Lafene Counseling and Psychological Services (CAPS). Academic support can be found with the Office of Student Life (OSL). OSL is a non-confidential resource. OIE also provides a comprehensive list of resources on their website. If you have questions about non-confidential and confidential resources, please contact OIE at equity@ksu.edu or (785) 532–6220. 









## Final Grades

A: 90+
B: 80–89.99
C: 70–79.99
D: 60–69.99
F: 0–59.99


*Tentative course schedule begins on next page*

\clearpage


## Tentative Course Schedule

```{r course calendar, dpi = 400, warning = FALSE, echo = FALSE, message = FALSE, fig.height=9, fig.width=7, fig.pos="center"}

library(ggplot2)
library(tidyverse)
# Create base data frame
dfr <- data.frame(date=seq(as.Date('2023-01-16'), as.Date('2023-05-31'), by = 1)) %>% 
  mutate(day = factor(strftime(date, format="%a"), levels=c("Sun","Mon","Tue","Wed","Thu","Fri","Sat")),
         week = factor(strftime(date, format="%U")),
         month = factor(strftime(date, format="%B"),levels=c("January","February","March", "April", "May")),
         ddate = factor(strftime(date, format="%d")),
         comment = case_when(
           date == as.Date("2023-02-03") ~ "Exam 1",
           date == as.Date("2023-03-10") ~ "Exam 2",
           date == as.Date("2023-04-14") ~ "Exam 3",
           date >= as.Date("2023-05-08") & date <=as.Date("2023-05-12") ~ "Exam 4 (Date TBD)",
           date >= as.Date("2023-03-13") & date <= as.Date("2023-03-17") ~ "Spring Break"
         ),
         week = as.numeric(week))



ggplot(dfr, aes(x = day, y = reorder(week, -week), fill = comment)) +
  geom_tile(color = "black", fill = "cornsilk1", size = 0.1) +
  geom_tile(color = "black", size = 0.1) +
  geom_text(aes(label = ddate), size = 5) +
  facet_wrap(. ~ month, scales = "free", ncol = 2) +
  #ghibli::scale_fill_ghibli_d("PonyoMedium") +
  viridis::scale_fill_viridis(discrete = TRUE, na.translate = FALSE) +
  theme_void() +
  theme(legend.position = "bottom",
        legend.direction = "vertical",
        legend.justification = "right",
        axis.text.x = element_text(size = 10, color = "black"),
        axis.text.y = element_text(size = 10, color = "white"),
        strip.text.x = element_text(size = 14, face = "bold", hjust = 0, margin = margin(b = 0.2, unit = "cm")),
        title = element_text(face = "bold"),
        panel.spacing = unit(0.8, "cm")) +
  guides(fill=guide_legend(ncol=1)) +
  labs(x = "Weekday",
       y = "Week #",
       fill = "Assignment")
```
  
  
  
\clearpage


## `r advdate(mon, 1)`: Course Introduction

### Topics:

 - Tuesday: Review the Syllabus. Discuss motivations for the course.
 - Thursday: The state in the modern era I
 
 - Read: Chapter 1 - What shaped our world?


## `r advdate(mon, 2)`: Background and Approach

### Topics: 

  - Tuesday: State and the modern era II
  - Thursday: Social scientific theories and methods I

  - Read: Chapter 2 - Understanding interests, interactions, and institutions

## `r advdate(mon, 3)`: Background and Approach

### Topics: 

  - Tuesday: Social scientific theories and methods II
  - Thursday: Slack day



## `r advdate(mon, 4)`: International Conflict

### Topics: 

  - Tuesday: War as an empirical phenomenon
  - Thursday: War as a bargaining process

  - Read: Chapter 3: Why are there wars?

## `r advdate(mon, 5)`: International Conflict

### Topics: 

  - Tuesday: Domestic politics and war
  - Thursday: Domestic politics and war

  - Read Chapter 4 - Domestic Politics and War

## `r advdate(mon, 6)`: International Institutions and Conflict

### Topics: 

  - Tuesday: Alliances
  - Thursday: Collective security
    - Guest Lecture: Lieutenant Commander Jean Anne Kim, USN.

  - Read: Chapter 5 - International Institutions and War

## `r advdate(mon, 7)`: Civil War and non-state actors

### Topics: 

  - Tuesday: Primer on civil wars
  - Thursday: Explaining rebel movements and civil conflict
  
  -  Read: Chapter 6 - Violence by nonstate Actors, pages 246--277


## `r advdate(mon, 8)`: Terrorism

### Topics: 

  - Tuesday: Primer on terrorism
  - Thursday: Understanding terrorism
  
  - Read: Chapter 6 - Violence by nonstate actors, pages 278--301


## `r advdate(mon, 9)`: Spring Break

### Topics: 

  - Tuesday: No class
  - Thursday: No class


## `r advdate(mon, 10)`: International Trade

### Topics: 

  - Tuesday: Thinking about international trade
  - Thursday: Theoretical models of international trade

  - Read: Chapter 7 - International Trade

## `r advdate(mon, 11)`: International Financial Relations

### Topics: 

  - Tuesday: Primer on international finance
  - Thursday: Theoretical models of international finance
  
  - Read: Chapter 8 - International Financial Relations

## `r advdate(mon, 12)`: International Monetary Relations

### Topics: 

  - Tuesday: Primer and mechanics of Monetary Policy
  - Thursday: In-class homework exercise on monetary policy
  
  - Read: Chapter 9 - International Monetary Relations



## `r advdate(mon, 13)`: International Development

### Topics: 

  - Tuesday: Primer on international development
  - Thursday: Understanding how other topics in IPE shape development
  
  - Read: Chapter 10: Causes of the wealth and poverty of nations


## `r advdate(mon, 14)`: Human Rights

### Topics: 

  - Tuesday: Historical background
  - Thursday: Trends in human rights conditions
  
  - Read: Chapter 12: Human rights


## `r advdate(mon, 15)`: Global Environmental Politics

### Topics: 

  - Tuesday: Basic issues in environmental politics
  - Thursday: Interests and cooperation in environmental politics
  
  -  Read: Chapter 13 - The global environment


## `r advdate(mon, 16)`: Challenges to the Global Order

### Topics: 

  - Tuesday: Emerging issues and challenges to global governance
  - Thursday: Review day



 

