---
title: "US Foreign Policy"
description: "Course syllabus"
date: "`r Sys.Date()`"
bibliography: ../../../classrefs.bib
csl: ../bib/apsr-bibliography.csl
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

mon <- as.Date("2022-08-22")

advdate <- function(obj, adv){
  tmon <- obj + 7*(adv - 1)
  tfri <- obj + 4 + 7*(adv - 1)
  tmon <- format(tmon, format = "%m/%d")
  tfri <- format(tfri, format = "%m/%d")
  zadv <- sprintf("%02d", adv)
  tmp <- paste("Week ", zadv, sep = '', ", ", tmon, " - ", tfri)
  return(tmp)
}

```
 

## Course Summary
 
This course focuses on the development of US foreign policy over the
course of the twentieth and twenty-first centuries. It will provide a brief background
on the history of US foreign policy before 1900, but will focus primarily on the developments and issues
occurring from 1900 through the present. The historical content of the course will be supplemented by
introducing students to political science literature that can help to identify and explain key events and trends in US foreign policy during this time period. 

## Course Objectives

1. The overarching goal is to develop an understanding
of the systematic forces that shape US foreign policymaking. Please note that this is different than
prescribing what policy *should* be (though this is also an important consideration when considering our motivations for studying the topic and reading course materials.
2. Identify the key actors that influence foreign policy.
3. Understand how and when the influence of actors varies (e.g. across time, issue area, etc.).
4. Identify the key issues confronting policymakers and society. 
5. Identify general trends in US foreign policy over time. 
    - What's changed? 
    - What stays the same?
    - When and why do we see changes in behavior or policies?
6. Students should develop a familiarity with social science research processes, practices. Specifically, students should have:
    - A basic understanding of how to read social scientific research articles. 
    - A basic understanding of research design considerations
    - A basic understanding of what the scope and limitations of research articles


## Course Format 

The course will be part lecture and part student participation. I encourage students to ask questions if the readings of lecture materials are unclear. In addition to class lectures students will be responsible for keeping up with readings and other class-related material as the semester progresses.


### Modality Changes

So what happens if we have to revert back to remote learning because of COVID? Ultimately my goal is to structure the course in a way to preserve our flexibility to the fullest extent possible. I have pre-recorded lectures available on my [YouTube channel](https://www.youtube.com/channel/UC7AzCLAufJAMCKkDL77clzw). In the event that we need to move to online only learning students can still view the lectures corresponding to each of the topics we're going to cover this semester. We will also use Zoom where needed to facilitate continued discussion. 

## Required Books 

There are three required books and several required articles (listed below in the course schedule).  You can view the full citation in the reference list at the end of hover your cursor over the citation below for a preview of the full title.

- @Allenetal2022
- @Karlin2022
- @GeorgeRishikof2017
- @McDowell2023


## Additional Readings

Students should also make an effort to stay informed on current events. Below are a some examples of good sources for keeping up with global events. Please note that some of these publications may be pay-walled, but you should have access through the university library.

  - [BBC News](www.bbc.com/news)
  - [The Monkey Cage](www.washingtonpost.com/news/monkeycage)
  - [Political Violence @ a Glance](www.politicalviolenceataglance.org)
  - [Foreign Affairs](www.foreignaffairs.com)
  - [Foreign Policy Magazine](www.foreignpolicy.com)
  - [The Conversation](https://theconversation.com/us)
  - [The Washington Post](www.washingtonpost.com)
  - [The New York Times](www.nytimes.com)
  - [The Economist](www.economist.com)
  - [The Financial Times](www.ft.com)
  

## Course Requirements

1. _Collaborative Discussion: 30%_

While the course has a heavy lecture component students will also spend time reflecting on class materials and discussing that material with classmates. We will use a combination of small group discussions and a Think-Pair-Share format---Students will have an opportunity to reflect on the course material independently, discuss the material with a classmate, and then communicate a summation of that discussion with the rest of the class. Students will then submit the notes from their discussion for review and grading. Accordingly, use the notes to keep track of the content of the discussion, ideas that emerge, and any lingering questions that might arise.


2. _Content Quizzes: 20%_

There will be 6-8 content quizzes throughout the semester and they will cover material we have already reviewed through readings, lectures, and discussion. I will announce these ahead of time. All quizzes are open book and open note assignments and will be completed through Canvas. I will also identify the appropriate readings corresponding to each quiz so you do not have to worry about guessing. 


3. _Photo Blog Journal: 20%_

Students will be required to keep a photo blog. You will be required to complete **3 posts** over the course of the semester. The goal of this assignment is to help students to develop a fuller appreciation for the ways in which their local community connects to the outside world. Here's the gist of the assignment:

  - Identify something around campus, in town, or in the nearby region that connects to US foreign policy and world politics. 
  - Take a picture of it!
  - Write blog post of 800-1000 words in which you:
    - Share the picture
    - Tell us what it is
    - Tell us where it is
    - Explain how it ties into US foreign policy and world politics
    - Share their posts with other students in the class
    - The post must include hyperlink citations or parenthetical references with full citation information where appropriate.
    
Remember, you must complete **3 posts** throughout the semester. You must also complete approximately **1 post per month!** These assignments require some thought, so plan accordingly. I will not accept late posts.
    
Be creative! There's really no right or wrong here, but it's on you to clearly explain how the subject of your photo ties into the course's subject matter. 
  
To complete this assignment you will have to create a blogger blog and submit the URL to Canvas. We will make sure this content is sharable through Canvas so students can explore each other's work. You can [watch a tutorial on how to set up your blog here](https://www.youtube.com/watch?v=xVy0_BcV7nU).
  
4. _The Music of Foreign Policy Wiki: 30%_

The arts are a powerful tool for communicating ideas across space and time, and are often deeply rooted in the politics of the moment. Over the last century music has increasingly come to serve as a vehicle for political speech, both reflecting and driving politics in the modern era. To that end, students will work together in small groups to build a Wiki-style page presenting a song or album that relates to US foreign policy in some way. The goal of this project is for students and their groups to build a wiki page on Canvas is dedicated to providing readers with a deeper understanding of the historical, political, and social underpinnings of songs. But there's a catch—since this is a class on US foreign policy, these songs have to be related to US foreign policy in some way. 

In general, each group's wiki page should focus on the following points:

  - *Song/Album*: Identify a song (or if you're feeling ambitious, an entire album) that speaks to some aspect of US foreign policy.
  - *Issue*: Identify the central issue, actors, policy, action, idea, behavior, etc., that is at the core of the song. 
  - *Context*: Once you have identified the central issue, provide some historical background information. This should help readers to better understand the social, political, and economic issues of the time period in question. Who are the key actors? What are the issues or policies of the time that serve as the focus for the song/album? 
  - *Reception*: How was the song/album received at the time of its release? Both critically and with respect to the popular impact that it had at the time?
  - *Legacy*: What is the legacy of the song/album? How does its legacy compare with its impact at the time of its release?


This will serve as your final project. Accordingly, I expect the final content to be detailed, polished, and thoroughly proofed. Your grade will be broken up into components and there will be multiple checkpoints along the way. The schedule is detailed below.

Deadlines: 

- September 9: Song/album choices due (2.5%)
- September 23: Final song/album selection due (2.5%)
- October 28: Annotated bibliography due (10%)
- November 18: Wiki draft due  (5%)
- December 9: Final Wiki page due (10%)
  
  
  
## Classroom Policies

1. Be kind. This might seem like an odd thing to list first, but I want to emphasize that this is not a normal class environment---this is still a time of unparalleled stress for many of us. We're still in the middle of a global pandemic, many of us are still very much worried about our own health and safety, and many of us are very much worried about the health and safety of our coworkers, friends, and families. Let's try to be cognizant of the fact that many of us (if not all) are going to be struggling at various points, and let's work together to make this semester as educational and enjoyable as possible. 

2. Be professional. We're all here to learn, and many of the issues that we discuss have real-world impacts that affect people we know, or in our broader community. We're here to work together to develop a better understanding of these issues. 

3. Study guides and extra credit. The class is the study guide. I do not offer extra credit.

4. Grade Appeals. If you believe that I have given you an incorrect grade on an exam, you may submit
a written appeal. All appeals must be in writing (they may not be made in person) and must contain
an explanation for why the grade is incorrect. I will also require you to wait 24 hours after receiving a
grade to submit a written appeal. Please note that even if your appeal is granted, I reserve the right
to regrade the assignment in its entirety, meaning that your grade may go up or down depending on
my reevaluation of the assignment.

5. The Syllabus. The syllabus is a living document that can and will be altered throughout the duration
of the course based both on need and design. Generally, this may mean readings will be removed or
added as needed. All changes will be listed on K-State Online, so make sure to check announcements
for any such change. Additionally, I maintain the right to all of my own intellectual property presented
in this course, whether it is the course lectures or this syllabus. Materials from this course ought not
to be reproduced without my expressed permission.

6. Classroom conduct. All student activities in the University, including this course, are governed by
the Student Judicial Conduct Code as outlined in the Student Governing Association By Laws, Article
V, Section 3, number 2. Students who engage in behavior that disrupts the learning environment may
be asked to leave the class.
  
  
## Academic Honesty and Plagiarism
  
All work turned in as a part of this class must be original.

Kansas State University has an Honor & Integrity System based on personal integrity which is presumed to be sufficient assurance in academic matters one’s work is performed honestly and without unauthorized assistance. Undergraduate and graduate students, by registration, acknowledge the jurisdiction of the Honor Integrity System. The policies and procedures of the Honor System apply to all full and part time
students enrolled in undergraduate and graduate courses on-campus, off-campus, and via distance
learning.

A component vital to the Honor & Integrity System is the inclusion of the Honor Pledge which applies to all
assignments, examinations, or other course work undertaken by students. The Honor Pledge is implied,
whether or not it is stated: “On my honor, as a student, I have neither given nor received unauthorized aid
on this academic work.”

The default in this class is that ALL work will be accomplished individually, UNLESS my permission is given
in advance of an assignment/quiz/exam/take-home exam/final. If you are in doubt, please ask.

A grade of XF can result from a breach of academic honesty. The F indicates failure in the course; the X
indicates the reason is an Honor Pledge violation.

For more information, visit the [Honor & Integrity System home web page](http://www.k-state.edu/honor/).
All work is expected to adhere to the standards outlined by the Kansas State University Handbook. All
written work is subject to online plagiarism checks and will be required for all reaction papers as well as the
final paper. Any cases of plagiarism will receive an automatic failure for the assignment, possible failure in
the course, and will be pursued further through the university for additional punishment.

The Kansas State University Handbook (2008, Appendix F) defines plagiarism as follows:

> Plagiarism is taking credit for someone else’s ideas, work, or words. In a university setting, it
means submitting academic, scholarly, or literary work in which you either claim or imply the
material to be your own, when that is not the case.

The Handbook also provides examples of plagiarism, and steps students can take to avoid it:

[P]lagiarism could include:

  - Buying a term paper, or copying another person’s paper, even if she or he gives you permission.
  - Cutting and pasting information from the internet into your own paper without properly
identifying it as a quotation and properly attributing the source.
  - Using someone else’s concept and presenting it as if it is your own original creativity, without
acknowledging the source of the idea.

While some acts of plagiarism are obvious attempts to deceive, like buying a term paper, others
result from sloppy scholarship or failure to follow proper format for crediting sources. For
example:

  - If you copy directly from another source and acknowledge it in your bibliography or list of
works cited, yet fail to put it in quotation marks, this is plagiarism. The reason is that,
although you have indicated that the idea is someone else’s, the lack of quotation marks
implies that the words are your own.
  - Paraphrasing is summarizing a source so that you have rewritten the material in your own
words but maintained the original author’s ideas. When you paraphrase, you do not need to
put the words in quotation marks, but you do need to properly attribute the original source.
Paraphrasing does not mean simply changing one or two words, or leaving out a sentence,
while the rest remains the same. Even if you identify the original author, but have used faulty
paraphrasing that retains too much of the original, the result is inadvertent plagiarism.
  - If you use another researcher’s specific methodology, you should acknowledge doing so.
While some research protocols are widespread and general enough to need no citation, if
the protocol, analysis, or technique can be attributed to a specific publication it should be
cited.
  - Works of visual art and design, and images of the work, cannot be misrepresented as
original work by another. Such images cannot be used in part or whole without permission
of the artist.

Also note that using papers previously written for another assignment in this, or in any other class, without
the express permission of the professor is grounds for failing an assignment. Using your own material in
this way, and not citing it, is also technically plagiarism and can be treated as such. If you have concerns
regarding any of these issues please do not hesitate to contact me.


## Statement on Mutual Respect and Inclusion in K-State Teaching and Learning Spaces

At K-State, faculty and staff are committed to creating and maintaining an inclusive and supportive learning environment for students from diverse backgrounds and perspectives. K-State courses, labs, and other virtual and physical learning spaces promote equitable opportunity to learn, participate, contribute, and succeed, regardless of age, race, color, ethnicity, nationality, genetic information, ancestry, disability, socioeconomic status, military or veteran status, immigration status, Indigenous identity, gender identity, gender expression, sexuality, religion, culture, as well as other social identities.

Faculty and staff are committed to promoting equity and believe the success of an inclusive learning environment relies on the participation, support, and understanding of all students. Students are encouraged to share their views and lived experiences as they relate to the course or their course experience, while recognizing they are doing so in a learning environment in which all are expected to engage with respect to honor the rights, safety, and dignity of others in keeping with the K-State Principles of Community https://www.k-state.edu/about/values/community/.

If you feel uncomfortable because of comments or behavior encountered in this class, you may bring it to the attention of your instructor, advisors, and/or mentors. If you have questions about how to proceed with a confidential process to resolve concerns, please contact the Student Ombudsperson Office. Violations of the student code of conduct can be reported here [https://www.k-state.edu/sga/judicial/student-code-of-conduct.html](https://www.k-state.edu/sga/judicial/student-code-of-conduct.html). If you experience bias or discrimination, it can be reported here [https://www.k-state.edu/report/discrimination/](https://www.k-state.edu/report/discrimination/).


## Statement Regarding Wearing of Face Coverings

Kansas State University strongly encourages, but does not require, that everyone wear masks while indoors on university property, including while attending in-person classes. For additional information and the latest on K-State’s face covering policy, see https://www.k-state.edu/covid-19/guidance/health/face-covering.html


## Statement Regarding Discrimination, Harassment, and Sexual Harassment

Kansas State University is committed to maintaining academic, housing, and work environments that are free of discrimination, harassment, and sexual harassment. Instructors support the University’s commitment by creating a safe learning environment during this course, free of conduct that would interfere with your academic opportunities. Instructors also have a duty to report any behavior they become aware of that potentially violates the University’s policy prohibiting discrimination, harassment, and sexual harassment (PPM 3010).  

If a student is subjected to discrimination, harassment, or sexual harassment, they are encouraged to make a non-confidential report to the University’s Office for Institutional Equity (OIE) using the online reporting form. Incident disclosure is not required to receive resources at K-State. Reports that include domestic and dating violence, sexual assault, or stalking, should be considered for reporting by the complainant to the Kansas State University Police Department or the Riley County Police Department. Reports made to law enforcement are separate from reports made to OIE. A complainant can choose to report to one or both entities. Confidential support and advocacy can be found with the K-State Center for Advocacy, Response, and Education (CARE). Confidential mental health services can be found with Lafene Counseling and Psychological Services (CAPS). Academic support can be found with the Office of Student Life (OSL). OSL is a non-confidential resource. A comprehensive list of resources is available here. If you have questions about non-confidential and confidential resources, please contact OIE at [equity@ksu.edu](mailto:equity@ksu.edu) or (785) 532–6220. 


## Citation and Style

All written work is expected to adhere to professional standards. As described above, citing previous work is
key. If you have any questions regarding the appropriate citation style please consult the American Political
Science Association style guide.


## Students with Disabilities

Students with disabilities who need classroom accommodations, access to technology, or information about
emergency building/campus evacuation processes should contact the Student Access Center and/or their
instructor. Services are available to students with a wide range of disabilities including, but not limited to,
physical disabilities, medical conditions, learning disabilities, attention deficit disorder, depression, and anxiety.
If you are a student enrolled in campus/online courses through the Manhattan or Olathe campuses,
contact the Student Access Center at [accesscenter@k-state.edu](accesscenter@k-state.edu), 785-532-6441; for Salina campus, contact
the Academic and Career Advising Center at acac@k-state.edu, 785-826-2649.

If you have any special conditions or needs that will affect your ability to complete the required assignments
outlined herein, please contact the instructor as soon as possible to discuss your needs. I am happy to help
you in any way that I can. With advance notice accommodations can be made prior to deadlines and tests.
Please note, however, that special consideration will not be granted afterwards as a means of addressing a
grade you do not like.



## Contacting the Instructor

The best way to get in touch with me is through email. Please do not hesitate to contact me if you have any questions about course content or your ability to complete the assigned material.



## Final Grades

A: 90+

B: 80–89.99

C: 70–79.99

D: 60–69.99

F: 0–59.99


*Tentative course schedule begins on next page*

\clearpage


## Tentative Course Schedule

```{r course calendar, dpi = 400, warning = FALSE, echo = FALSE, message = FALSE, fig.height=9, fig.width=7, fig.pos="center"}

pacman::p_load(
  "ggplot2",
  "showtext"
)

# Add custom fonts
#sysfonts::font_add_google("Roboto", family = "oswald")
#showtext::showtext_auto()

# Create base data frame
dfr <- data.frame(date=seq(as.Date('2022-08-01'), as.Date('2022-12-31'), by = 1))
dfr$day <- factor(strftime(dfr$date,format="%a"), levels=c("Sun","Mon","Tue","Wed","Thu","Fri","Sat"))
dfr$week <- factor(strftime(dfr$date,format="%U"))
dfr$month <- factor(strftime(dfr$date,format="%B"),levels=c("August","September","October", "November", "December"))
dfr$ddate <- factor(strftime(dfr$date,format="%d"))

# Label dates
dfr$comment <- "Available"
dfr$comment[dfr$date < as.Date("2022-08-22") | dfr$date>as.Date("2021-12-09")] <- NA
dfr$comment[dfr$date==as.Date("2022-09-09")] <- "Song/album choice due"
dfr$comment[dfr$date==as.Date("2022-09-23")] <- "Final song/album choice due"
dfr$comment[dfr$date==as.Date("2022-10-28")] <- "Annotated bibliography due"
dfr$comment[dfr$date==as.Date("2022-11-18")] <- "Wiki draft due"
dfr$comment[dfr$date==as.Date("2022-12-09")] <- "Final Wiki due"

# Photo blogs
dfr$comment[dfr$date==as.Date("2022-09-30")] <- "Blog post due"
dfr$comment[dfr$date==as.Date("2022-10-31")] <- "Blog post due"
dfr$comment[dfr$date==as.Date("2022-11-30")] <- "Blog post due"


# Thanksgiving
dfr$comment[dfr$date >= as.Date("2022-11-21") & dfr$date <= as.Date("2022-11-25")] <- "Thanksgiving Break"

dfr$week <- as.numeric(dfr$week) - min(as.numeric(dfr$week) - 1)
dfr$comment <- factor(dfr$comment, levels=c("Blog post due", "Song/album choice due", "Final song/album choice due", "Annotated bibliography due", "Wiki draft due", "Thanksgiving Break", "Final Wiki due"))

ggplot(dfr, aes(x = day, y = reorder(week, -week), fill = comment)) +
  geom_tile(color = "black", size = 0.1) +
  geom_text(aes(label = ddate), size = 3) +
  facet_wrap(. ~ month, scales = "free", ncol = 2) +
  theme_void() +
  theme(legend.position = "bottom",
        legend.direction = "vertical",
        legend.justification = "right",
        axis.text.x = element_text(size = 10, color = "black"),
        axis.text.y = element_text(size = 10, color = "white"),
        strip.text.x = element_text(size = 13, face = "bold", hjust = 0),
        strip.text = element_text(size = 16, face = "bold", margin = margin(b = 0.2, unit = "cm")),
        title = element_text(face = "bold")) +
  guides(fill=guide_legend(ncol=2)) +
  scale_fill_brewer(palette = "Set2", na.translate = FALSE) +
  labs(x = "Weekday",
       y = "Week #",
       fill = "Assignment")
```
  
  
  
\clearpage


## `r advdate(mon, 1)`: Course Introduction

1. Watch video on [how to read a social science research paper](https://www.youtube.com/watch?v=pZq6c_shco0)
2. Set up your blog!


## `r advdate(mon, 2)`: Historical background

### Readings and Assignments: 

- Read @fordham2019



## `r advdate(mon, 3)`: The Cold War and US Hegemony

### Readings and Assignments:

- Read @NSC68



## `r advdate(mon, 4)`: The Cold War and Domestic Politics

### Readings and Assignments:

- Read @dudziak2004
- Read @Fordham2007


## `r advdate(mon, 5)`: The Presidency and Congress

### Readings and Assignments:

- Read @GeorgeRishikof2017[Chapter 13]
- Read @lupton2017


## `r advdate(mon, 6)`: The Policymaking Process

### Readings and Assignments:

- Read @GeorgeRishikof2017[Chapters 1 and 2]
- Read @miller2016



## `r advdate(mon, 7)`: The Bureaucracy / The State Department

### Readings and Assignments:

- Read @keane2015
- Read @GeorgeRishikof2017[Chapter 4]



## `r advdate(mon, 8)`: The Defense Department and the Military

### Readings and Assignments:

- Read @Karlin2022



## `r advdate(mon, 9)`: The Use of Military Force

### Readings and Assignments:

- Watch [Restrepo](https://watchdocumentaries.com/restrepo/)
- Read @sigar2019



## `r advdate(mon, 10)`: Security Cooperation

#### Guest Lecture: 
- Lieutenant Commander Jean Anne Kim, USN on US alliance relations in the Asia-Pacific region.
- Friday, October 28

### Readings and Assignments:

- Read @beckley2015
- Read [Flynn, Allen, and Martinez Machain (2019)](https://theconversation.com/why-does-the-us-pay-so-much-for-the-defense-of-its-allies-5-questions-answered-127683)
- Read @moon2009


## `r advdate(mon, 11)`: Trade, Immigration, Finance, and Monetary Policy

### Readings and Assignments:

- Read @Oneil2022
- Read @rosellacappellazielinski2022


## `r advdate(mon, 12)`: Trade, Immigration, Finance, and Monetary Policy (cont.)

### Readings and Assignments:

- Read @McDowell2023



## `r advdate(mon, 13)`: Foreign Aid

- Read @milner2013


## `r advdate(mon, 14)`: Thanksgiving Break (No Class)

### Readings and Assignments:

1. Get some rest
2. Enjoy yourselves
3. Eat good food
4. Take a good nap


## `r advdate(mon, 15)`: Human Rights Issues

### Readings and Assignments:

- Read @Power2001 


## `r advdate(mon, 16)`: Multilateralism and global governance

### Readings and Assignments:

- Read @schultz2017
- Read @busby2020 





 

