---
title: "Social Science Methods Primer Notes"
author: Michael Flynn
---

# Background

This lecture is going to serve as a basic primer to help students understand the role of theory and statistical methodology.

1. Theory
    - Models
    - Variables
    - Actors
    - Interests
    - Institutions
2. Statistics
    - Sampling
    - Confidence intervals
    - Substantive importance


# Theory

- "A logically consistent set of statements that explains a phenomenon of interest." FLS, Glossary A-7
- Theory is the story that we tell about how things work. 
- It describes the relationships between two or more variables
- It can be used to predict outcomes

## Models

- Models are based on theory. They are simplistic representations of the phenomena we are trying to explain or understand. 
- Models can vary in complexity but are necessarily simplifications because the real world is too complex to replicate perfectly. 
- We often want to understand lots of different phenomena, so moving between models is also important.

## Variables

- What do we mean by variables?
- Variables: The basic elements or moving parts of our model, the value of which changes across time or space.
- I.e., they necessarily vary.


## Actors

- Actors are the people in our models who "do" things. 
- They are the players in the game


## Interests

- The things actors want
- The goals they pursue


## Institutions

- The rules of the game
- Can be formal or informal. Laws, norms, or other rules that structure how actors relate to one another and how processes unfold.
- Note this is different than institutions in a purely organizational sense.

# Statistics

Statistics is one tool that we use in the social sciences, but it is possible to conduct research without it. But this class will read a lot of material that uses stats so let's get comfortable with some basics. 


## Sampling

- This is one of the most important aspects of statistical work
- We often don't have data on, or can't observe, the full population of cases.
- The limited portion we observe and/or have data on is called the sample (think of an urn full of balls)
- Public opinion surveys are another good example

- Small samples present problems and may not be representative of the population about which we're interested in drawing inferences.

```{r, warning=FALSE, message=FALSE}
library(tidyverse)

# Set random number generator seed to ensure replicability
set.seed(66502)

# Simulate some data for population
voters <- tibble(party = c(rep("Democrat", 105000), rep("Republican", 95000)))

# Check the relative counts
table(voters$party)

# Check the percentage 
percent <- sum(grepl("Democrat", voters$party))/length(voters$party)
print(percent)

# Draw one from the population
sample.small <- sample(voters$party, size = 1)
print(sample.small)


# Let's gather a slightly larger sample
sample.medium <- sample(voters$party, size = 20)
percent.medium <- sum(grepl("Democrat", sample.medium))/length(sample.medium)
print(percent.medium)


# Now let's get a BIG sample
sample.large <- sample(voters$party, size = 1000)
percent.large <- sum(grepl("Democrat", sample.large))/length(sample.large)
print(percent.large)


# Now let's do that a lot!
samplelist <- tibble(mean = rep(NA, 1000)) |> 
  rowwise() |> 
  mutate(mean = (sum(grepl("Democrat", sample(voters$party, size = 1000)))/1000))

# Show distribution of sample means
hist(samplelist$mean, breaks = 30)

# Print the mean of the sampling means
print(mean(samplelist$mean))

```


# Statistical significance and substantive significance.

- Not an indicator of substantive importance
- Technical definition: You take the ratio of the coefficient to the standard error to derive a $z$ or $t$ score and then use that score to position the coefficient/ratio on a normal or Student's t distribution.
- The p value is the probability of getting a ratio as big as or larger than the observed one if the data were generated at random. 


```{r}
set.seed(123)
x <- seq(1, 100, 1)
error <- rnorm(100, 0, 0.5)
y = 0 + 0.005 * x + error

data <- tibble(x = x,
               y = y)


ggplot(data, aes(x = x, y = y)) +
  geom_point() +
  geom_smooth(method = "lm") + 
  geom_hline(yintercept = 0, size = 1.5, color = "red")
```
